---
title: "Lab 3"
author: "Diana Tang"
format: 
  html:
    embed-resources: true
editor: visual
---

## Libraries

```{r}
library(tidyverse)
library(data.table)
```

Step 1: **Read in the data**

```{r}
download.file(
  "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
  destfile = file.path("~", "Downloads", "met_all.gz"),
  method   = "libcurl",
  timeout  = 60
)

met <- data.table::fread(file.path("~", "Downloads", "met_all.gz"))
```

Step 2: **Check the dimensions**

```{r}
dim(met)
```

There are **2,377,343 rows** and **30 columns** in the met dataset.

```{r}
tail(met)
```

Step 3: **Take a look at the variables**

```{r}
str(met)
```

Step 4: **Take a closer look at the key variables.**

```{r}
table(met$year)
```

```{r}
table(met$day)
```

```{r}
table(met$hour)
```

```{r}
summary(met$temp)
```

```{r}
summary(met$wind.sp)
```

```{r}
summary(met$elev)
```

```{r}
met[met$elev==9999.0] <- NA
summary(met$elev)
```

At what elevation is the highest weather station? **It is at 4113 meters.**

```{r}
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)
```

Step 5: **Check the data against an external data source.**

```{r}
met <- met[temp>-15]
met2 <- met[order(temp)]
head(met2)
```

**Here we removed temperatures colder than -15C. The new dataset has minimum temp -3C which is reasonable.**

Step 6: **Calculate summary statistics**

```{r}
elev <- met[elev==max(elev), ]
summary(elev)
```

```{r}
cor(elev$temp, elev$wind.sp, use="complete")
```

```{r}
cor(elev$temp, elev$hour, use="complete")
```

```{r}
cor(elev$wind.sp, elev$day, use="complete")
```

```{r}
cor(elev$wind.sp, elev$hour, use="complete")
```

```{r}
cor(elev$temp, elev$day, use="complete")
```

Step 7: **Exploratory graphs**

```{r}
hist(met$elev, breaks=100)
```

```{r}
hist(met$temp)
```

```{r}
hist(met$wind.sp)
```

**The highest weather station is located close to Denver, CO.**

```{r}
library(leaflet)
```

```{r}
leaflet(elev) |>
  addProviderTiles('OpenStreetMap') |>
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)
```

```{r}
library(lubridate)
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)
```

```{r}
elev <- elev[order(date)]
head(elev)
```

```{r}
plot(elev$date, elev$temp, type='l')
```

```{r}
plot(elev$date, elev$wind.sp, type='l')
```

Summarizing trends: Between August - September, wind elevation ranged between 0 - 20 MPH. There was an oscillatory behavior exhibited in the plots. The wind elevation in itself seemed to be trending upwards from August to September.

```{r}
met_stations <- (unique(met[,c("lat","lon")]))  
dim(met_stations)
```

```{r}

leaflet(met_stations) |> 
  addProviderTiles('CartoDB.Positron') |> 
  addCircles(lat = ~lat, lng = ~lon, opacity = 1, fillOpacity = 1, radius = 400)
```

**Step 8: Ask questions**

```{r}
table(met$dew.point)
summary(met$dew.point)
```

```{r}
hist(met$dew.point, breaks = 100)
```

```{r}
plot(elev$date, elev$dew.point, type = 'l')
```

How does dew point vary with date from August through September? It varies through an oscillatory behavior with a maximum of 6-7 and minimum of -6. It appears to vary in an oscillatory behavior over a time span of 1 day.
